---
title: "covid19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

## Calling libs

```{r lib}
library(tidyverse)
library(tidylog)
library(plotly)
```


## Reading data
```{r data}
d <- read_csv("data/processed.csv")
```


## Gulf countries


```{r gulf data}
g_countries <- c("Saudi Arabia", "United Arab Emirates", "Bahrain")

d_gulf <- 
  d %>% 
  filter(country_region %in% g_countries,
         date == max(date))
```

Saudi Arabia has the highest number of death cases but the number is reasonable given the number of people living in the kingdom. 

```{r}
 d_gulf %>% 
  ggplot(aes(x = reorder(country_region, -deaths_num), y = deaths_num))+
  geom_col()+
  labs(
    x = "",
    y = "number of death cases",
    title = "Saudi Arabia has the height est death cases among gulf countries"
  )
```

## Highest confirmed and deaths countries

China has the highest recovered cases. 
```{r}

highest_confirmed_countries <- 
  d %>% 
  filter(date == max(date)) %>% 
  pivot_longer(-c(date, country_region), names_to = "case_type") %>% 
  filter(case_type == "confirmed_num") %>% 
  arrange(desc(value)) %>% 
  head(10) %>% 
  pull(country_region)

highest_deaths_countries <- 
   d %>% 
  filter(date == max(date)) %>% 
  pivot_longer(-c(date, country_region), names_to = "case_type") %>% 
  filter(case_type == "deaths_num") %>% 
  arrange(desc(value)) %>% 
  head(10) %>% 
  pull(country_region)


highest_recovered_countries <- 
   d %>% 
  filter(date == max(date)) %>% 
  pivot_longer(-c(date, country_region), names_to = "case_type") %>% 
  filter(case_type == "recovered_num") %>% 
  arrange(desc(value)) %>% 
  head(10) %>% 
  pull(country_region)
  
d_top_countries <- 
  c(highest_recovered_countries, highest_deaths_countries, highest_recovered_countries) %>% 
  unique()


```


```{r}
d %>% 
  filter(
    date == max(date), 
    country_region %in% d_top_countries
    ) %>% 
 pivot_longer(-c(date, country_region), names_to = "case_type") %>% 
  ggplot(aes(x = reorder(country_region, value), y = value, fill= case_type))+
  geom_col(position = "dodge")+
  coord_flip()+
  labs(
    x = "",
    y = "",
    title = "China has the highest recovery numbers"
  )
```

## number of deaths in Saudi Arabia compared to top 10 % countries


```{r}
deaths_top_countreis <- 
  d %>% 
  filter(
    date == max(date),
    ) %>% 
  select(-confirmed_num, - recovered_num) %>% 
  arrange(desc(deaths_num)) %>% 
  head(10)

saudi <-
   d %>% 
  filter(
    date == max(date),
    country_region == "Saudi Arabia"
    ) %>% 
  select(-confirmed_num, -recovered_num)


deaths_top_countreis <-
  deaths_top_countreis %>% 
  rbind(saudi)


deaths_top_countreis %>% 
  ggplot(aes(x = reorder(country_region, deaths_num), y = deaths_num))+
  geom_col()+
  coord_flip()+
  scale_y_continuous(labels = scales::comma)+
  labs(
    y = "number of death cases",
    x = "",
    title = "Saudi Arabia has the lowset number of death cases compared with other countries"
  )
  

```


## comparing Saudi Arabia to countries have the same range of number of deaths

Saudi Arabia has the second highest number of recovered cases. 
```{r}
d %>% 
  filter(
    date == max(date),
    confirmed_num < 1650 & confirmed_num > 1300
    ) %>% 
  arrange(desc(confirmed_num)) %>% 
  pivot_longer(-c(date, country_region), names_to = "case_type") %>% 
  ggplot(aes(x = reorder(country_region, value), y = value, fill = case_type))+
  geom_col(position = "dodge")
```


```{r}
similar_countries <- 
  d %>% 
  filter(
    date == max(date),
    confirmed_num < 1650 & confirmed_num > 1300
    ) %>% 
  pull(country_region)


```

```{r}
d %>% 
  filter(country_region %in% similar_countries) %>% 
  ggplot(aes(x = date, y = confirmed_num, col= country_region))+
  geom_line()
ggplotly()
```

